The problem and promise of scale in multilayer networks

### 2020-02-03 paper/manuscript.Rmd
update indiv fitness perspective

Social network positions are frequently used to represent the social environment
that an individual experiences or their social phenotype. Social phenotypes are
increasingly linked to important fitness for individuals of various species,
including primates (e.g. Brent, Ruiz-Lambides, & Platt, 2017; Thompson, 2019),
dolphins (e.g. Stanton & Mann, 2012; Stanton, Gibson, & Mann, 2011), and birds
(e.g. McDonald, 2007; Royle, Pike, Heeb, Richner, & Kolliker, 2012). The effect
of scale on the relationship between an individual’s social network position and
fitness may depend on the hypotheses being tested and logistical constraints.
Two scales that are important for understanding fitness consequences in the
context of social network positions are temporal and social scales (Almeling,
Hammerschmidt, Sennhenn-Reulen, Freund, & Fischer, 2016; Berger, Lemaître,
Allainé, Gaillard, & Cohas, 2015; Brent et al., 2017).

Social network structure depends on the timeframe of observation, which can make
networks difficult to compare (Castles et al., 2014). Furthermore, social
network structure can vary seasonally and indicate social environments that an
individual experiences (e.g. Brent, MacLarnon, Platt, & Semple, 2013; Hamede,
Bashford, McCallum, & Jones, 2009; Holekamp, Smith, Strelioff, Van Horn, &
Watts, 2012), or they can be consistent across time, which can be used to
reflect social phenotypes (Aplin et al., 2015; Stanley, Mettke-Hofmann, Hager, &
Shultz, 2018). Seasonal fluctuations have the potential to have disease
transmission and fitness effects for individuals (Balasubramaniam et al., 2019;
Chen et al., 2014). <!-- **[QW: I'm not sure what this paragraph adds - I think if we
need to save space we could cut it]**-->

In long-lived species, social phenotypes can change over their lifetimes in ways
that can affect fitness (Berger et al., 2015; Brent et al., 2017). These
findings highlight the necessity to carefully consider the temporal scale that
is meaningful for the proposed hypotheses, especially for understanding the
relationship between an individuals social phenotype and fitness. Further
complicating the issue, social phenotypes can change over stages of ontogeny,
although it is important to note that individuals progress through development
at different rates (Tarka, Guenther, Niemelä, Nakagawa, & Noble, 2018). This
variation can make defining appropriate temporal scales for testing the fitness
implications of an individual’s social phenotype over different stages of
ontogeny even more difficult. However, considering fine temporal scales may
provide insight across development stages that may otherwise be missed
(CITE-myself??). Multilayer networks provide a framework to test hypotheses
linking social phenotypes and fitness across temporal scales. Because of the
non-independent nature of social interactions and associations, especially over
time, it is hard to test hypotheses about social dynamics with a suitable null
control (Farine, 2017; Proskurnikov & Tempo, 2017). Using the framework of
multilayer networks, making layers of different temporal periods provides a way
to make appropriate null models to test questions about the dynamics of an
individual’s social network position over multiple aggregated time periods.

Social scale is another major consideration for social network analysis,
especially in the context of different types of interactions and associations in
the same analysis (Carter, Lee, & Marshall, 2015; Castles et al., 2014; Farine,
2015). It is a common trope that individuals must be associating to interact and
most studies therefore assume that proximity is proxy for interacting (Farine,
2015). For instance, baboons need to be in proximity in order to groom each
other, but this trope highlights human bias toward interactions that can be
readily observed in the field. Animals have certain interactions that do not
require proximity such as long-distance vocalization and olfactory signals
(Carter et al., 2015). With the advancement of bio-logging technology, we have
the ability to record social interactions that do not require proximity,
particularly long-distance communication networks with devices like microphone
arrays (Snijders & Naguib, 2017). The multilayer network framework further
enables researchers to examine the relationship between social phenotypes and
fitness across social scales and to test how good social proximity is as a proxy
for social interactions in a holistic way.

These different social scales can have different fitness repercussions for
individuals. Rhesus macaques (*Macaca mulatta*) that had stronger, more stable
social associates and grooming partners, but not more social partners, had
higher survivorship than those who did not (Ellis, Snyder-Mackler,
Ruiz-Lambides, Platt, & Brent, 2019). This finding highlights the need to
investigate the effect of temporal and social scales on the relationship between
social network positions and fitness. Multilayer networks provide an avenue to
do so considering the whole social system rather than just one type of social
interaction at a time, which is what has been done under most circumstances but
is not very realistic.

Multilayer networks further provide unique challenges and opportunities for
understanding the social phenotypes and environments of individuals and their
fitness outcomes. Picking the proper temporal scale is important but a challenge
because all layers in the multilayer network should reflect the same time
period, and the period should be long enough for each layer to have enough data
for robust networks (Farine, 2017). These considerations may be limiting factors
when answering questions that require timeframes that are relevant to different
individuals. On the other hand, multilayer network metrics provide an
opportunity to simplify statistical models. Interactions that happen meters
apart, proximity associations, and long-distance interactions can be
encapsulated in a single multilayer network, and a single metric can be used to
describe the various social scales of an individual’s network position. This
multilevel network metric could be used to assess the relationship between an
individual's social phenotype and fitness. It would also be a meaningful
simplification of many network metrics that could help simplify statistical
analyses and deal with the frequent correlations between social network metrics
in different types of networks (Castles et al., 2014).

>>>>

Social network positions are frequently used to represent the social environment
that an individual experiences or their social phenotype. Social phenotypes are
increasingly linked to important fitness for individuals of various species,
including primates (e.g. Brent, Ruiz-Lambides, & Platt, 2017; Thompson, 2019),
dolphins (e.g. Stanton & Mann, 2012; Stanton, Gibson, & Mann, 2011), and birds
(e.g. McDonald, 2007; Royle, Pike, Heeb, Richner, & Kolliker, 2012). Both
temporal, seasons and stages of life, and social scales, types of interactions,
have emerged as important for understanding fitness consequences in the context
of social network positions (Almeling, Hammerschmidt, Sennhenn-Reulen, Freund, &
Fischer, 2016; Berger, Lemaître, Allainé, Gaillard, & Cohas, 2015; Brent,
Ruiz-Lambides, & Platt, 2017; Holekamp, Smith, Strelioff, Van Horn, & Watts,
2012). Multilayer networks provide a way to bridge these different scales so
that we can examine social phenotypes in more holistic ways rather than looking
at each scale separately as has traditionally been done.

For instance, social scale is major consideration for social network analysis,
especially in the context of different types of relationships, such as grooming
and association, that are typically examined separately while answering a single
question (Carter, Lee, & Marshall, 2015; Castles et al., 2014; Farine, 2015).
Furthermore, it is a common trope that individuals must be associating to
interact, and many studies therefore assume that proximity is proxy for
interacting (Farine, 2015). For instance, baboons need to be in proximity in
order to groom each other. However, animals have certain interactions that do
not require proximity such as long-distance vocalization and olfactory signals
(Carter et al., 2015). With the advancement of biologging technology, we have
the ability to record social interactions that do not require proximity,
particularly long-distance communication networks with devices like microphone
arrays (Snijders & Naguib, 2017). These varying social scales can have different
fitness repercussions for individuals. Rhesus macaques (Macaca mulatta) had
higher survivorship when they had fewer but stronger, more stable social
associates and grooming partners rather than simply more associates (Ellis,
Snyder-Mackler, Ruiz-Lambides, Platt, & Brent, 2019). This finding highlights
that different qualities of social scales have lasting consequences.

Temporal scale is also important for understanding the repercussions of social
phenotypes. For instance, social phenotypes can change over ontogeny in ways
that can affect fitness (Berger et al., 2015; Brent et al., 2017). Thus, it is
critical to carefully consider the temporal scale that is meaningful for the
proposed hypotheses. Further complicating the issue of selecting suitable time
scales across ontogeny, individuals progress through development at different
rates (Tarka, Guenther, Niemelä, Nakagawa, & Noble, 2018). This variation can
make defining appropriate temporal scales for testing the fitness implications
of an individual’s social phenotype over different stages of ontogeny even more
difficult when they vary by individual. However, considering fine temporal
scales provide insight across developmental stages that may otherwise be missed
(Turner, Bills, & Holekamp, 2018).

Multilayer networks provide unique opportunities and challenges, linking these
various scales, for understanding the social phenotypes and environments of
individuals and their fitness outcomes. The multilayer network framework enables
researchers to examine these different social scales simultaneously rather than
individually, which is what has been done under most circumstances, but is not
very realistic. Metrics from multilayer networks provide an opportunity to
simplify statistical models. Interactions that happen side-by-side, proximity
associations, and long-distance interactions can be encapsulated in a single
multilayer network, and a single metric can be used to describe the various
social scales of an individual’s network position. This multilevel network
metric could be used to assess the relationship between an individual’s social
phenotype and fitness. It would also be a meaningful simplification of many
network metrics that could help simplify statistical analyses and deal with the
frequent correlations between social network metrics in different types of
networks (Castles et al., 2014). Furthermore, multilayer networks provide a
framework to test hypotheses linking social phenotypes and fitness across
temporal scales. Because of the non-independent nature of social interactions
and associations, especially over time, it is hard to test hypotheses about
social dynamics with a suitable null control (Farine, 2017; Proskurnikov &
Tempo, 2017). Using the framework of multilayer networks, making layers of
different temporal periods provides a way to make appropriate null models to
test questions about the dynamics of an individual’s social network position
over multiple aggregated time periods.

On the other hand, picking the proper temporal scale is important but a
challenge because all layers in the multilayer network should reflect the same
time period, and the period should be long enough for each layer to have enough
data for robust networks (Farine, 2017). These considerations may be limiting
factors when answering questions that require timeframes that are relevant to
different individuals, for instance. In both mono- and multilayer networks, the
effect of scale on the relationship between an individual’s social network
position and fitness need to be considered based on the hypotheses being tested
and logistical constraints.
